---
title: "Plots for Superconductor Dataset"
author: "Mariya Kim"
date: "10/8/2021"
output:
  pdf_document: default
  
---
 
Plots for all **Superconductor Dataset** variables were additionally used for data analysis
to select plots for the report and discover relationships and correlations.   

(1) Distributions of the variables.  Histograms of all variables show, that there 
no normally distributed variables in this data set and some have low variance.  

(2) *critical_temp* vs variables, colored by *number_of_elements*. There are variables 
that can be stratified by *number_of_elements* and can be found using this set of graphs.  

(3) Correlations in the groups of predictors.  

(4) Cluster heatmaps of feature groups also show that patterns can be found across
all plots; some features are paired:  
  &nbsp;&nbsp;&nbsp;   *maen* and *gmean*;  
&nbsp;&nbsp;&nbsp;   *std* and *wtd_std*;  
&nbsp;&nbsp;&nbsp;   *entropy* and *wtd_entropy*;  
&nbsp;&nbsp;&nbsp;   *wtd_mean* and *wtd_gmean.*  
  indicating similar predictors that were obtained using similar calculations.  
&nbsp;&nbsp;&nbsp;  Also *_Density* variables stand out in many cluster heatmaps,
as well as "range_" features.

```{r, warning=FALSE, message=FALSE}
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(readxl)) install.packages("readxl")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(caret)) install.packages("caret")
if(!require(tidyr)) install.packages("tidyr")
if(!require(reshape2)) install.packages("reshape2")
if(!require(tinytex)) install.packages("tinytex")
if(!require(dplyr)) install.packages("dplyr")
if(!require(stringr)) install.packages("stringr")
if(!require(RColorBrewer)) install.packages("RColorBrewer")
if(!require(matrixStats)) install.packages("matrixStats")
if(!require(tinytex)) install.packages("tinytex")
if(!require(knitr)) install.packages("knitr")

```



```{r, warning=FALSE, message=FALSE}
fileName <- "4_Superconductors.xlsx"
superconductors <- read_excel(fileName)
set.seed(1, sample.kind="Rounding") # if using R 3.5 or earlier, use `set.seed(1)`
index_val <- createDataPartition(y = superconductors$critical_temp,
                                 times = 1, p = 0.2, list = FALSE)
# sc set, that does not have validation set rows.
sc <- superconductors[-index_val, ] 
# 25%  of sc set is used for Critical_temp-vs-variables plots.
sc_small <- slice_sample(sc, prop = 0.25)
```

\pagebreak

### (1) Distributions of all variables in **Superconductivty** data set.
```{r, message=FALSE, fig.show="hold", out.width="35%"}
cols <- 1:82
plot_histogram_function <- sapply(cols, function(cols){
  df <- sc[cols] %>% as.matrix()
  plot <- hist(df, 
               xlab=" ",
               main = paste("Histogram of",colnames(sc)[cols]))
  plot
})
```

\pagebreak

### (2) Critical temperature vs variables, colored by *number_of_elements*.
```{r, message=FALSE, fig.show="hold", out.width="35%"}
cols <- 2:81
plot_function <- sapply(cols, function(cols){
  df <- cbind(sc_small[cols], sc_small[82])
  plot(df, col=sc_small$number_of_elements)
})
```

\pagebreak

### (3) Correlations in the groups of predictors.
```{r, warning=FALSE, message=FALSE}
col_name <- str_subset(colnames(sc), "atomic_mass") # length(col_name)
patterns1 <- str_replace(col_name, "_atomic_mass", "")
patt_unite <- data.frame(a = rep("^", 10), b = patterns1) 
patterns1 <- unite(patt_unite, a, b, col = "c", sep = "") %>% pull()
string <- toString(patterns1)
string <- str_replace_all(string, ", ", "|")
patterns2 <- str_replace(colnames(sc), string, "") %>% unique() 
patterns2 <- patterns2[2:9] 
p <- c(patterns2, patterns1) 
```

Correlation heatmaps of features that are grouped using patterns:
```{r}
p
```


```{r, fig.show="hold", message=FALSE,  out.width="35%"}
corr_heatmaps <- map(p, function(p){
  corr_group <-  sc[str_subset(colnames(sc), p)] %>%
    cor() %>% data.frame()
  
  corr_group <- rownames_to_column(corr_group, "col_name") 
  corr_group <- melt(corr_group) %>% 
    set_names(c("col_name", "variable", "value"))
  
  plot <- corr_group %>%
    ggplot(aes(col_name, variable, fill = value)) +
    geom_tile() +
    scale_fill_distiller(palette = "Reds", direction = 1)  +
    labs(caption =  paste("Correlation heatmap of the variables in", p, "group.")) +
    theme_minimal()+
    theme(plot.caption.position = "panel", 
          plot.caption = element_text(size = 14, face = "bold.italic", hjust = 0.65),
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 90, hjust = 1))
  plot
})

corr_heatmaps
```
   
   \pagebreak
   
### (4) discovering clusters in the groups of features using heatmap() function.  
Features are grouped using patterns:
```{r}
p
```

```{r, fig.show="hold", message=FALSE, warning=FALSE, out.width="35%"}
x <- sc[1:81]  %>% as.matrix()
y <- sc[82] %>% round() %>% pull()
rownames(x) <- y   # critical_temp column is turned into row names
x <- sweep(x, 2, colMeans(x)) %>% as.matrix() # removing the center

heatm_clust_col <- sapply(p, function(p){
  ind <- which(colnames(sc) %in% colnames(sc[str_subset(colnames(sc), p)]))
  heatmap(t(x[,ind]), col = brewer.pal(11, "Spectral"), 
          scale = "column") 
})

heatm_clust_col
```
   
